A Spiking Independent Accumulator Model for Winner-Take-All

### 2017-01-24 doc/paper/paper.tex
Change equation reference.
Edit IA model description.

We implement Equation~\ref{eqn:usher-mcclelland} with the NEF by using one population of neurons for each $x_i$, and applying Principle~3 to each population.
By appropriately selecting the gain and bias parameters from Principle~1, we ensure that each state variable is rectified ($x_i \ge 0$) as required.
We believe this implementation is novel as it does not interpret each $x_i$ as a distinct neural firing rate, but rather as a population-level representation distributed across any number of spiking neurons.
In effect, heterogeneous spike trains are weighted by optimal decoding weights to precisely implement the stated dynamics.
%As such, the model equations do not impose restrictions on the neuron model or requires to lump the neurons into a single population firing rate.
This allows us to attain greater biological realism without altering the dynamics prescribed by (\ref{eqn:usher-mcclelland}).
\todo{distributing the states across neurons would be yet another implementation; maybe worth mentioning in the discussion...as that might be even more bio-plausible}

@@

\subsection{Independent accumulator model}
The other WTA mechanism that we investigate is our proposed independent accumulator (IA) model.
We use the term `independent' to refer to the fact that there are no direct interactions between each accumulator, unlike in the LCA model which has direct competition between states.
To enable a form of competition, we add a second thresholding layer that  projects back to self-excite and mutually-inhibit the first layer.
We now provide the details of each layer.
\todo{Network diagram?}

The first layer consists of a separate integrator population (i.e., accumulator) for each state variable $x_i(t), \ i < D$.
The second layer consists of independent, non-recurrent populations that receive input from the first layer in a one-to-one fashion.
From the second layer, we decode the function $\bar{x}_i := \Theta(x_i - \vartheta)$ where $\Theta$ is the Heaviside function, and $\vartheta = 0.8$ is a fixed threshold value that determines how much evidence needs to be accumulated to produce an output.
The Heaviside decoded output of layer 2 projects back to layer 1 to add $\bar{x}_i - \bar{\beta} \sum_{j \neq i} \bar{x}_j$ to the input of $x_i$.
Since intuitively the largest input will accumulate the fastest, once this reaches the threshold $\vartheta$ it will self-excite and inhibit all other state variables.
Fixing $\bar{\beta} = 2$ ensures that the losing state variables will go to zero (see supplementary analysis).
This is summarized more precisely by the following dynamics (see Fig.~\ref{fig:indacc}):
\begin{equation}
    \begin{split}
        \frac{{\partial x}_i}{\partial t} = \lambda \rho_i \frac{1}{\tau_1} + \left( 
            \bar{x}_i - \bar{\beta} \sum_{j \neq i} \bar{x}_j \right) \frac{1}{\tau_2} , \quad x_i &\ge 0 .
    \end{split}
\end{equation}
We set $\lambda = 1$ to be consistent with the LCA model.
% should we mention that we use Principle 3 to implement each integrator
% should we say anything about how the thresholding population works?

>>>>

We implement Equation~\ref{eqn:usher-mcclelland} with the NEF by using one population of neurons for each $x_i$, and applying Principle~3 to each population.
By appropriately selecting the gain and bias parameters from Principle~1, we ensure that each state variable is rectified ($x_i \ge 0$) as required.
We believe this implementation is novel as it does not interpret each $x_i$ as a distinct neural firing rate, but rather as a population-level representation distributed across any number of spiking neurons.
In effect, heterogeneous spike trains are weighted by optimal decoding weights to precisely implement the stated dynamics.
%As such, the model equations do not impose restrictions on the neuron model or requires to lump the neurons into a single population firing rate.
This allows us to attain greater biological realism without altering the dynamics prescribed by Equation~\ref{eqn:usher-mcclelland}.
\todo{distributing the states across neurons would be yet another implementation; maybe worth mentioning in the discussion...as that might be even more bio-plausible}

@@

\subsection{Independent accumulator model}
The other WTA mechanism that we investigate is our proposed independent accumulator (IA) model.
We use the term `independent' to refer to the fact that there are no direct interactions between each accumulator, unlike in the LCA model which has direct competition between states.
To enable a form of competition, we add a second thresholding layer that  projects back to self-excite and mutually-inhibit the first layer.
We now provide the details of each layer, again implemented with the principles of the NEF.
\todo{Network diagram?}

The first layer consists of a separate integrator population (i.e., accumulator) for each state variable $x_i(t), \ i < D$.
The second layer consists of independent, non-recurrent populations that receive input from the first layer in a one-to-one fashion.
From the second layer, we decode the function $\bar{x}_i := \Theta(x_i - \vartheta)$ where $\Theta$ is the Heaviside function, and $\vartheta = 0.8$ is a fixed threshold value that determines how much evidence needs to be accumulated to produce an output.
The Heaviside decoded output of layer 2 projects back to layer 1 to add $\bar{x}_i - \bar{\beta} \sum_{j \neq i} \bar{x}_j$ to the input of $x_i$.
Since intuitively the largest input will accumulate the fastest, once this reaches the threshold $\vartheta$ it will self-excite and inhibit all other state variables.
Fixing $\bar{\beta} = 2$ ensures that the losing state variables will go to zero (see supplementary analysis).
This is summarized more precisely by the following dynamics (see Fig.~\ref{fig:indacc}):
\begin{equation}
    \begin{split}
        \frac{{\partial x}_i}{\partial t} = \rho_i \frac{1}{\tau_1} + \left( 
            \bar{x}_i - \bar{\beta} \sum_{j \neq i} \bar{x}_j \right) \frac{1}{\tau_2} , \quad x_i &\ge 0 .
    \end{split}
\end{equation}
% should we say anything about how the thresholding population works?
